---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Data-Management-Plans

## Useful Field of View Study Data
Esin Gürcan and M.D. Rutherford
Department of Psychology, Neuroscience and Behavior, McMaster University, Hamilton, On, Canada

_Correspondence_ concerning this paper should be addressed to M.D. Rutherford, Department of Psychology, Neuroscience and Behavior, McMaster University, 1280 Main Street West, Hamilton, Ontario, Canada, L8S 4K1
_Corresponding author e-mail_: rutherm@mcmaster.ca
_First author e-mail_:grcane@mcmaster.ca

*_Abstract_*
  Research has shown that autistic individuals have difficulties in broadening peripheral vision (Mann & Walker, 2003) and in divided attention (McDougal et al., 2020; Boxhoorn et al., 2018). On the other hand, the Broader Autism Phenotype (BAP) refers to sub-clinical autistic traits that are more common in relatives of individuals with autism compared to the general population (Rubenstein & Chawla, 2018; Hartley et al., 2019), and these divided attention difficulties are also shown in the siblings of autistic individuals (Belmonte et al., 2010). However, Rutherford et al. (2009)’s study using the Useful Field of View (UFOV) paradigm showed advantages in divided attention for young autistic adults compared to a control group. 

  In the present project, we replicated Rutherford et al.'s (2009) UFOV paradigm, which involves extracting central letter identification and localizing peripheral targets (at eccentricities of 4, 8, 12, 16, and 20 degrees) under focused and divided attention tasks. In the divided attention task, both targets were presented, whereas the focused attention task required participants to identify only one type of target. We conducted this project with parents and family members of individuals with autism (as BAP) to explore whether a UFOV advantage exists among autism families at later ages, given that the error rate in UFOV performance tends to increase with age (Sekuler et al., 2000).
  
  The current sample (data collection in progress) consists of 12 parents with a familial history of autism (Mean age = 39.58 years, SD = 6.14; 3 male, 1 gender-fluid, 1 non-binary, 7 female) and 9 parents without such a history (Mean age = 38.89 years, SD = 2.47; 5 male, 4 female). Participants were assessed using the WAIS-IV, the Autism Quotient (AQ) (cut-off score = 26) (Baron-Cohen et al., 2001), and a shortened version of the UFOV paradigm. The order of the assessments was cross-randomized. Error rate and reaction time in task performance were analyzed based on task type (central and peripheral) using an ANOVA: 2 (group: AQ cut-off above vs. AQ cut-off below OR familial history: yes vs. no) x 2 (attention: focused vs. divided) x 5 (eccentricity: 4, 8, 12, 16, 20).

*_Keywords:_* divided attention, autism, adults, broader autism phenotype, genetics

## Specification Table
Specification | Description 
----------| ------
Subject | * Developmental Psychology, * Experimental and Cognitive Psychology
Specific subject area | Autism and Broader Autism Phenotype: We look at the genetic component of autism through conducting the studies with their family members, focusing on potential perceptual strengtsh assosicated with autism
Type of Data | Raw, processed and analyzed data with tables
Data collection |   In the present project, we replicated Rutherford et al.'s (2009) UFOV paradigm, which involves extracting central letter identification and localizing peripheral targets (at eccentricities of 4, 8, 12, 16, and 20 degrees) under focused and divided attention tasks. In the divided attention task, both targets were presented, whereas the focused attention task required participants to identify only one type of target. We conducted this project with parents and family members of individuals with autism (as BAP) to explore whether a UFOV advantage exists among autism families at later ages, given that the error rate in UFOV performance tends to increase with age (Sekuler et al., 2000). Participants will fill out two questionnaires and be tested with the Wechsler Adult Intelligence Scale (WAIS). The WAIS is used to assess general cognitive ability in adults; however, the test results will not be shared with them One questionnaire focuses on their demographic background, while the other is the Autism Quotient (AQ) questionnaire. The AQ questionnaire allows individuals to self-assess the extent to which they exhibit characteristics associated with autism. After that, the actual experiment session will begin.   The experimental session includes three tasks. In the first task, they will be asked to identify the letters that they have seen as targets, and feedback will be provided based on their correct or incorrect answers. In the second task, they will see a filled circle with spokes and be asked to identify which spoke it is in. In the last task, they will see both letters and a filled circle and be asked to identify both the letters and which spoke the circle is in.Feedback will be provided when they see the letters. They will be asked to complete these tasks. This study lasts approximately 1 hour (including the WAIS) and 30 minutes (including the experiment and filling out the forms). They can withdraw your participation at any time without any consequences. The error proportions from the experimental data were adjusted using the inverse sine of their square root to normalize the variance of our measurements. Exclusion criteria were having a diagnosis of autism, and not having a normal-to corrected vision/hearing.
Data source location | Data collection was conducted in the lab environment.
Data accessibility | All raw and processed data will be added to the Open Science Framework, along with a data dictionary and the analysis code. (I will add the DOI and link once the study is completed.)
Related research article | 1. Rutherford, M. D., Richards, E. D., Moldes, V., & Sekuler, A. B. (2007). Evidence of a divided-attention advantage in autism. Cognitive Neuropsychology, 24(5), 505–515. https://doi.org/10.1080/02643290701508224 2. Sekuler, A.,B., Bennett, P. J., Mamelak, A.M., (2000). Effects of Aging on the Useful Field of View. Experimental Aging Research, 26(2), 103–120. https://doi.org/10.1080/036107300243588

## Value of the data
_Importance of the Data_

* This data is particularly important for our purpose of highlighting potential perceptual strengths associated with autism and potentially corroborating that autism is not purely negative; instead, there are strengths associated with it. 
* This builds on the study by Rutherford et al. (2007) and aims to replicate the results in an older age group, as individuals' perceptual skills tend to decrease with age.

_Benefit_

* Autism communities can benefit from our results.

_Reusability_

* This data is reusable through our shared data. Especially in meta-analysis.


## Data Description

* Raw data can be found in raw_data.csv.
* Processed data can be found in processed_data.csv.
* Participant demographics can be found in demographics.csv.
* Data cleaning and analysis code can be found in UFOV_Analysis.Rmd.
* ANOVA tables can be found in the folder named ANOVA Results.
* Correlation tables can be found in the folder named Correlation Results.

## Experimental Design, Materials and Methods
Data were acquired through in-person assessments and experimental software. In the in-person assessments, the WAIS-IV was administered by undergraduate and graduate students. Participants also filled out a form for demographic information and the Autism Quotient. Then, they responded behaviorally to the stimuli in the experiment created using PsychoPy software. The data analysis and processing were conducted using R Studio software. The R Markdown can be found in the OSF folder. The data cleaning and analysis procedures followed the guidelines based on Rutherford et al. (2007)'s study.

## Declarations
* We hereby confirm that there is no conflict of interest.
* This study was approved by the McMaster Research Ethics Board (MREB), and participants provided consent at each assessment.
* CRediT author statement: 
  * EG: Data collection, Data analysis, Writing, Original draft preparation
  * MDR: Conceptualization, Methodology, Supervision

## Acknowledgements
  We would like to express our gratitude to the families who participated in this study. We also extend our thanks to the undergraduate research assistants for their invaluable help with data collection. In particular, we acknowledge, William Clark, Fatemeh Ilkhani, Reem Mokahal, Rupinder Pamma.

### References
1. Baron-Cohen, S., Wheelwright, S., Skinner, R., Martin, J., & Clubley, E. (2001). The autism-spectrum quotient (AQ): evidence from Asperger syndrome/high-functioning autism, males and females, scientists and mathematicians. Journal of autism and developmental disorders, 31(1), 5–17. https://doi.org/10.1023/a:1005653411471
2. Belmonte, M. K., Gomot, M., & Baron‐Cohen, S. (2010). Visual attention in autism families: ‘unaffected’ sibs share atypical frontal activation. Journal of Child Psychology and Psychiatry and Allied Disciplines., 51(3), 259–276. https://doi.org/10.1111/j.1469-7610.2009.02153.x
3. Boxhoorn, S., Lopez, E., Schmidt, C., Schulze, D., Hänig, S., & Freitag, C. M. (2018). Attention profiles in autism spectrum disorder and subtypes of attention-deficit/hyperactivity disorder. European Child & Adolescent Psychiatry, 27(11), 1433–1447. https://doi.org/10.1007/s00787-018-1138-8
4. Hartley, S. L., Hickey, E. J., DaWalt, L., & Rodriguez, G. (2019). Broader autism phenotype and couple interactions in parents of children with autism. Autism, 23(8), 2068–2079. https://doi.org/10.1177/1362361319841312 
5. Mann, T. A., & Walker, P. (2003). Autism and a deficit in broadening the spread of visual attention. Journal of Child Psychology and Psychiatry, 44(2), 274–284. https://doi.org/10.1111/1469-7610.00120
6. McDougal, E., Riby, D. M., & Hanley, M. (2020). Profiles of academic achievement and attention in children with and without Autism Spectrum Disorder. Research in Developmental Disabilities, 106, 103749–103749. https://doi.org/10.1016/j.ridd.2020.103749
7. Rubenstein, E., & Chawla, D. (2018). Broader Autism Phenotype in Parents of Children with Autism: A Systematic Review of Percentage Estimates. Journal of Child and Family Studies, 27(6), 1705–1720. https://doi.org/10.1007/s10826-018-1026-3  
8. Rutherford, M. D., Richards, E. D., Moldes, V., & Sekuler, A. B. (2007). Evidence of a divided-attention advantage in autism. Cognitive Neuropsychology, 24(5), 505–515. https://doi.org/10.1080/02643290701508224   
9. Sekuler, A.,B., Bennett, P. J., Mamelak, A.M., (2000). Effects of Aging on the Useful Field of View. Experimental Aging Research, 26(2), 103–120. https://doi.org/10.1080/036107300243588


```{r}
library(usethis)
usethis::use_git()
usethis::use_github()
```
